A comparison of design-based and model-based approaches for spatial data.

### 2021-10-10 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
updated text surrounding the confidence interval figure.

colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
effplot <- ggplot(data = combo_data, aes(x = n_factor, y = rel_efficiency,
                              colour = approach)) +
  facet_grid(psill_ratio ~ resptype) +
  geom_jitter(width = 0.19, size = 2.5) +
  scale_colour_manual(values = colour_scale) +
  theme_bw(base_size = 14) +
  labs(x = "Sample Size", colour = "Approach",
       y = "Relative Efficiency")
effplot
```

@@

MAJOR POINTS for the following Figure:

- coverage for the model-based estimator is slightly higher than coverage for the design-based estimator in the normal settings. Coverages are about equal in the lognormal settings with a slight edge to model-based (this is the point that will be tougher to explain: I would have expected the design-based estimator to have better coverage in the lognormal settings because it has fewer assumptions).
- coverage is at or near the nominal 95% in all of the normal settings, where assumptions for the model approach and the design approach are satisfied.
- for the model-based approach, the more skewed the population is, the higher the sample size needed to satisfy CLT for predicting a mean. The derivation of the BLUP is entirely moment-based (no distribution assumed) but we still need to assume a distribution to estimate spatial parameters and to generate bounds of a prediction interval.
- many confidence intervals generated for design-based approaches also rely on the CLT and the normal distribution to generate the interval. Again, for highly skewed data with a small sample size, this assumption is violated even though all of the assumptions for generating the estimator are valid.

```{r}
ggplot(data = combo_data, aes(x = n_factor, y = coverage, colour = approach)) +
  geom_hline(yintercept = 0.95) +
  geom_jitter(width = 0.19, size = 2.5) + 
  facet_grid(psill_ratio ~ resptype) +
  scale_colour_manual(values = colour_scale) +
  ylim(0.80, 1) +
  theme_bw(base_size = 14) +
  labs(x = "Sample Size", colour = "Approach",
       y = "Interval Coverage")
```

@@

>>>>

colour_scale <- c("goldenrod1", "goldenrod4", "mediumpurple1", "mediumpurple4")
effplot <- ggplot(data = combo_data, aes(x = n_factor, y = rel_efficiency,
                              colour = approach)) +
  facet_grid(psill_ratio ~ resptype) +
  geom_jitter(width = 0.24, size = 2.5) +
  scale_colour_manual(values = colour_scale) +
  theme_bw(base_size = 14) +
  labs(x = "Sample Size", colour = "Approach",
       y = "Relative Efficiency")
effplot
```

@@

Figure \ref{fig:figconf} shows the coverage for each of the four approaches. We see that the four approaches have somewhat similar coverages in all settings, with GRTS-Design having slightly lower coverage when the response is normal.

In the normal response settings, where assumptions for the model-based approaches and the design-based approaches are satisfied, all approaches have coverage around the nominal 0.95. Because the intervals are symmetric, normal-based intervals, the coverages are generally lower than the nominal 95%. We see that, because the sampling distribution of the mean is asymptotically normal (reference for this? a spatial version of CLT), the coverages when the sample size is 200 are much closer to the nominal 95% than the coverages when the sample size is 50. In general, the more skewed the distribution of the response, the larger the sample size needed to ensure proper coverage of these normal-based intervals.

Do we want to mention stuff like what's in these next bullet points or no?

- for the model-based approach, the more skewed the population is, the higher the sample size needed to satisfy CLT for predicting a mean. The derivation of the BLUP is entirely moment-based (no distribution assumed) but we still need to assume a distribution to estimate spatial parameters and to generate bounds of a prediction interval.
- many confidence intervals generated for design-based approaches also rely on the CLT and the normal distribution to generate the interval. Again, for highly skewed data with a small sample size, this assumption is violated even though all of the assumptions for generating the estimator are valid.

```{r figconf, out.width = "100%", fig.cap = "Coverage of the four design-analysis approaches. All confidence intervals are normal-based and have a nominal confidence level of 0.95, marked with a horizontal line. The plot is faceted by the type of response on the columns and the partial-sill to total-variance ratio on the rows.", message = FALSE, warning = FALSE}
ggplot(data = combo_data, aes(x = n_factor, y = coverage, colour = approach)) +
  geom_hline(yintercept = 0.95) +
  geom_jitter(width = 0.24, size = 2.5) + 
  facet_grid(psill_ratio ~ resptype) +
  scale_colour_manual(values = colour_scale) +
  ylim(0.80, 1) +
  theme_bw(base_size = 14) +
  labs(x = "Sample Size", colour = "Approach",
       y = "Interval Coverage")
```

@@

<!-- not sure why this is getting cutoff. temporary fix: -->

\pagebreak

### 2021-10-10 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
updated text surrounding the confidence interval figure.

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\usepackage{longtable}
\usepackage{graphicx}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={A comparison of design-based and model-based approaches for spatial data.},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

MAJOR POINTS for the following Figure:

\begin{itemize}
\tightlist
\item
  coverage for the model-based estimator is slightly higher than
  coverage for the design-based estimator in the normal settings.
  Coverages are about equal in the lognormal settings with a slight edge
  to model-based (this is the point that will be tougher to explain: I
  would have expected the design-based estimator to have better coverage
  in the lognormal settings because it has fewer assumptions).
\item
  coverage is at or near the nominal 95\% in all of the normal settings,
  where assumptions for the model approach and the design approach are
  satisfied.
\item
  for the model-based approach, the more skewed the population is, the
  higher the sample size needed to satisfy CLT for predicting a mean.
  The derivation of the BLUP is entirely moment-based (no distribution
  assumed) but we still need to assume a distribution to estimate
  spatial parameters and to generate bounds of a prediction interval.
\item
  many confidence intervals generated for design-based approaches also
  rely on the CLT and the normal distribution to generate the interval.
  Again, for highly skewed data with a small sample size, this
  assumption is violated even though all of the assumptions for
  generating the estimator are valid.
\end{itemize}

\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}

@@

>>>>

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\usepackage{longtable}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={A comparison of design-based and model-based approaches for spatial data.},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

Figure \ref{fig:figconf} shows the coverage for each of the four
approaches. We see that the four approaches have somewhat similar
coverages in all settings, with GRTS-Design having slightly lower
coverage when the response is normal.

In the normal response settings, where assumptions for the model-based
approaches and the design-based approaches are satisfied, all approaches
have coverage around the nominal 0.95. Because the intervals are
symmetric, normal-based intervals, the coverages are generally lower
than the nominal 95\%. We see that, because the sampling distribution of
the mean is asymptotically normal (reference for this? a spatial version
of CLT), the coverages when the sample size is 200 are much closer to
the nominal 95\% than the coverages when the sample size is 50. In
general, the more skewed the distribution of the response, the larger
the sample size needed to ensure proper coverage of these normal-based
intervals.

Do we want to mention stuff like what's in these next bullet points or
no?

\begin{itemize}
\tightlist
\item
  for the model-based approach, the more skewed the population is, the
  higher the sample size needed to satisfy CLT for predicting a mean.
  The derivation of the BLUP is entirely moment-based (no distribution
  assumed) but we still need to assume a distribution to estimate
  spatial parameters and to generate bounds of a prediction interval.
\item
  many confidence intervals generated for design-based approaches also
  rely on the CLT and the normal distribution to generate the interval.
  Again, for highly skewed data with a small sample size, this
  assumption is violated even though all of the assumptions for
  generating the estimator are valid.
\end{itemize}

\begin{figure}
\includegraphics[width=1\linewidth]{SpatialDVM_Manuscript_files/figure-latex/figconf-1} \caption{Coverage of the four design-analysis approaches. All confidence intervals are normal-based and have a nominal confidence level of 0.95, marked with a horizontal line. The plot is faceted by the type of response on the columns and the partial-sill to total-variance ratio on the rows.}\label{fig:figconf}
\end{figure}

@@

\pagebreak
